---
title: "When R Environments Matter"
output:
  html_document:
    toc: true
    number_sections: false
    toc_float:
      collapsed: true
      smooth_scroll: true
---

Saturday, 10 December 2016

<br>

# Intro

Which environment you are using in R matters, but we generally ignore this fact. Here is a pretty simple example of when the environment does matter in a context where such problems often arise—where we define an object that will be called by a function.

# The two-function problem

The main function (ending in "a") takes three arguments: `x, y, z` and feeds them to a second function (ending in "b"). This "b" function only accepts `x` as an argument, but the function uses `y` and `z`, in addition to `x` (not ideal function writing).

<br>

## Option 1: Non-nested functions

The first pair of functions (1a and 1b) are not nested, meaning `test1b` is not defined within `test1a`.


```{r, error = TRUE}
# Define the first function:
# This function has three arguments and calls the next function
test1a <- function(x, y, z){
    test1b(x)
}
# Define the second function:
# This function uses x, y, and z, but it only takes x as an argument
test1b <- function(x){
    print(rep(x, y * z))
}
# Run the function 'test1a'
test1a(1, 2, 3)
```

And we get an error: R cannot find `y` (or `z`). This error occurs because the function `test1b` lives on a different level (environment) than the environment in which we defined `x, y, z`. The three variables are defined within the context of the function `test1a`, but `test1b` is defined outside of `test1a` and only has access to the variable `x`, which we pass it through its `x` argument.

One solution: nest the functions...

<br>

## Option 2: Nested functions

Simply define the second (b) function inside of the first (a) function.

```{r, error = TRUE}
# Define the first function:
# This function has three arguments and calls the next function
test2a <- function(x, y, z){
  # Define the second function:
  # This function uses x, y, and z, but it only takes x as an argument
  test2b <- function(x){
      print(rep(x, y * z))
  }
  test2b(x)
}
# Run the function 'test2a'
test2a(1, 2, 3)
```

It works (no errors).

<br>

## Option 3: Write clearer functions

There other ways to fix the error in Option 1. The one you would probably hear from StackOverflow is to write better functions (_i.e._ name all the arguments the function needs). Errors within user-defined functions often come from users (me) getting lazy and calling objects within the function that have not been passed to the function. This tendency is a bit sloppy and, as Option 1 demonstrates, error prone.

The fix: pass all three arguments from the first (a) function to the second (b) function.

```{r, error = TRUE}
# Define the first function:
# This function has three arguments and calls the next function
test3a <- function(x, y, z){
    test3b(x, y, z)
}
# Define the second function:
# This function uses x, y, and z, but it only takes x as an argument
test3b <- function(x, y, z){
    print(rep(x, y * z))
}
# Run the function 'test3a'
test3a(1, 2, 3)
```

Success. Plus, you won't get reprimanded for sloppy scripting.

<br>

# Wrap up

Environments matter. Sloppy coding can remind you how they matter. Loops also will remind you how they matter (for instance, `lapply` vs. for loops while updating data.tables).

Enjoy!
